---
title: "ps3"
output: html_document
---

```{r}
# packages
library(readxl)
library(lubridate)
library(ggplot2)
```

# read data
```{r}
master <- read_excel("master.xlsx")
```

# create new variables dataframe for master.xlsx
```{r}
CA_post <- data.frame( date= format(as.Date(ym(paste(master$year,master$month))), "%y-%m"),
                       s1= ifelse(master$month==1,1,ifelse(master$month==12,-1,0)),
                       s2= ifelse(master$month==2,1,ifelse(master$month==12,-1,0)),
                       s3= ifelse(master$month==3,1,ifelse(master$month==12,-1,0)),
                       s4= ifelse(master$month==4,1,ifelse(master$month==12,-1,0)),
                       s5= ifelse(master$month==5,1,ifelse(master$month==12,-1,0)),
                       s6= ifelse(master$month==6,1,ifelse(master$month==12,-1,0)),
                       s7= ifelse(master$month==7,1,ifelse(master$month==12,-1,0)),
                       s8= ifelse(master$month==8,1,ifelse(master$month==12,-1,0)),
                       s9= ifelse(master$month==9,1,ifelse(master$month==12,-1,0)),
                       s10= ifelse(master$month==10,1,ifelse(master$month==12,-1,0)),
                       s11= ifelse(master$month==11,1,ifelse(master$month==12,-1,0)),
                       pi_mnppr= master$pi_mnppr,
                       price= master$all_ad_p,
                       ret_m_al= master$ret_m_al,
                       trend= master$trend,
                       cpi= master$cpi,
                       quantity= master$all_ad_v,
                       log_quantity= log(master$all_ad_v))
```

# summary master.xlsx data
```{r}
summary(CA_post)
```

# replicate demand regression
```{r}
org_model <- lm(log(quantity) ~ s1+s2+s3+s4+s5+s6+s7+s8+s9+s10+s11+log(pi_mnppr/cpi)+log(price/cpi)+log(ret_m_al/cpi)+log(trend), data=CA_post)
summary(org_model)
```



# predict demand based on original model
```{r}
org_pred <- predict(org_model, newdata = CA_post)

#comparison 
plot_a <- ggplot(CA_post, aes(date, log_quantity, group=1)) +
  geom_point(aes(color = "Actual Demand"), size = 1) +
  geom_line(aes(color = "Actual Demand")) +
  geom_point(aes(y = org_pred, color = "Predicted Demand of Original Model"), size = 1) +
  geom_line(aes(y = org_pred, color = "Predicted Demand of Original Model")) +
  scale_color_manual(values = c("red", "blue")) +
  scale_x_discrete(breaks = CA_post$date[c(T,F,F)]) +
  theme(axis.text.x=element_text(size=4,angle = 90))

plot_a + labs(x = "Date", y = "Demand") 
```


# build higher order demand regression
```{r}
ho_model <- lm(log(quantity) ~ s1+s2+s3+s4+s5+s6+s7+s8+s9+s10+s11+trend+I(log(trend)^2)+I(trend^2), data=CA_post)
summary(ho_model)
```

# predict demand based on original model
```{r}
ho_pred <- predict(ho_model, newdata = CA_post)

#plot two predictions and actual demand
plot_b <- ggplot(CA_post, aes(date, log_quantity, group=1)) +
  geom_point(aes(color = "Actual Demand"), size = 1) +
  geom_line(aes(color = "Actual Demand")) +
  geom_point(aes(y = org_pred, color = "Predicted Demand of Original Model"), size = 1) +
  geom_line(aes(y = org_pred, color = "Predicted Demand of Original Model")) +
  geom_point(aes(y = ho_pred, color = "Predicted Demand of Higher Order Model"), size = 1) +
  geom_line(aes(y = ho_pred, color = "Predicted Demand of Higher Order Model")) +
  scale_color_manual(values = c("red", "green", "blue")) +
  scale_x_discrete(breaks = CA_post$date[c(T,F,F)]) +
  theme(axis.text.x=element_text(size=4,angle = 90))

plot_b + labs(x = "Date", y = "Demand") 
```


# create new variables dataframe for newmaster.xlsx
```{r}
new_master <- read_excel("newmaster.xlsx")

CA_post_new <- data.frame( date= format(as.Date(ym(paste(new_master$year,new_master$month))), "%y-%m"),
                       s1= ifelse(new_master$month==1,1,ifelse(new_master$month==12,-1,0)),
                       s2= ifelse(new_master$month==2,1,ifelse(new_master$month==12,-1,0)),
                       s3= ifelse(new_master$month==3,1,ifelse(new_master$month==12,-1,0)),
                       s4= ifelse(new_master$month==4,1,ifelse(new_master$month==12,-1,0)),
                       s5= ifelse(new_master$month==5,1,ifelse(new_master$month==12,-1,0)),
                       s6= ifelse(new_master$month==6,1,ifelse(new_master$month==12,-1,0)),
                       s7= ifelse(new_master$month==7,1,ifelse(new_master$month==12,-1,0)),
                       s8= ifelse(new_master$month==8,1,ifelse(new_master$month==12,-1,0)),
                       s9= ifelse(new_master$month==9,1,ifelse(new_master$month==12,-1,0)),
                       s10= ifelse(new_master$month==10,1,ifelse(new_master$month==12,-1,0)),
                       s11= ifelse(new_master$month==11,1,ifelse(new_master$month==12,-1,0)),
                       pi_mnppr= new_master$pi_mnppr,
                       price= new_master$all_ad_p,
                       ret_m_al= new_master$ret_m_al,
                       trend= new_master$trend,
                       cpi= new_master$cpi,
                       log_quantity= log(new_master$all_ad_v))

new_period <- CA_post_new[CA_post_new$date %in%
                                c("96-06","96-07","96-08","96-09","96-10","96-11","96-12","97-01","97-02","97-03","97-04","97-05","97-06","97-07"),]
```

# summary newmaster.xlsx data
```{r}
summary(new_period)
```

# predict new demand based on higher order regression
```{r}
org_pred_np <- predict(org_model, newdata = new_period)
ho_pred_np <- predict(ho_model, newdata = new_period)
```

# plot two predictions and actual demand
```{r}
plot_c <- ggplot(new_period, aes(date, log_quantity, group=1)) +
  geom_point(aes(color = "Actual Demand"), size = 1) +
  geom_line(aes(color = "Actual Demand")) +
  geom_point(aes(y = org_pred_np, color = "Predicted Demand of Original Model"), size = 1) +
  geom_line(aes(y = org_pred_np, color = "Predicted Demand of Original Model")) +
  geom_point(aes(y = ho_pred_np, color = "Predicted Demand of Higher Order"), size = 1) +
  geom_line(aes(y = ho_pred_np, color = "Predicted Demand of Higher Order")) +
  scale_color_manual(values = c("red", "green", "blue")) +
  theme(axis.text.x=element_text(size=10,angle = 90))

plot_c + labs(x = "Date", y = "Demand") 
```



 
